### 2015-12-31 article.rmd
Pushed more details and contents

---
title: "InnoDB Full text Search hacking with Shiny/R"
author: "Emanuel Calvo - Pythian"
date: "December 22, 2015"
output: html_document
runtime: shiny
---

```{r init, message = FALSE, echo=FALSE}
library(png)
#options(width = 9999)
#knitr::opts_chunk$set(fig.width=9999)
# opts_chunk$set(fig.width=9999) #, fig.height=6))

library(knitr)
library(shiny)
require(RMySQL)
library(ggplot2)

# Comment if you want to avoid reload all the data
# If you want to start from scratch, uncomment the following line.
# Otherwise, you can restore the dumps located in the dumps folder
#source('~/fts_article/load.R')

# Yeah, only for development environements
all_cons <- dbListConnections(MySQL())
for(con in all_cons) dbDisconnect(con)

# You do not want clear passwords, but is a sandbox.
#con <- dbConnect(MySQL(),
#                 user="msandbox", password="msandbox", port=14901
#                 ,dbname="test", socket="/tmp/mysql_sandbox14901.sock",
#                 host="127.0.0.1")

con <- dbConnect(MySQL(),
                 user="ftsinnodb", password="elfilete", port=3306
                 ,dbname="ftsinnodb", #socket="/tmp/mysql_sandbox14901.sock",
                 host="db4free.net")

```

## Version

 **BETA - NOT FINAL**

This R Markdown document is made interactive using Shiny. Unlike the more traditional
workflow of creating static reports, you can now create documents that allow your
readers to change the assumptions underlying your analysis and see the results immediately.  
To learn more, see [Interactive Documents](http://rmarkdown.rstudio.com/authoring_shiny.html).

Repository available at [Github](https://github.com/3manuek/fts_article).

## Credits

Author: Emanuel Calvo
Company: Pythian

Thanks to Valerie Parham-Thompson @ Pythian and Daniel Prince @ Oracle.

Thanks db4free.net for the MySQL free online database!

## Requirements

If you want to run this locally, please install a MySQL 5.7 instance and:

- Uncomment the line __source('~/fts_article/load.R')__ and run the report, or
- Restore the dumps located in the `dumps` folder. Uncomment the table creation.

## Some initial thoughts

A couple of days ago one of our customers came up with a question regarding FTS
over InnoDB engine. Although the question is not answered in the current article,
I came up with the conclusion that FTS is sometimes misunderstood.

The point of this "article" is to show dynamically how the search algorithms work,
using non-fictional data (data sources were downloaded from www.gutenberg.org
project) within an easy interface.

In order to show the effects off the field sizes over the query expansion algorithm,
you will see two main tables (bookContent and bookContentByLine) both containing
the same books in different approaches: by line and by paragraph. You'll see the
noise generated by the `QUERY EXPANSION` algorithm when phrases are too large.

The current article has been developed using Shiny/R in order to allow you to see
the effects of the algorithms. It connects to a open database at `db4free.net`,
which unfortunately is not 100% available.

## Action!

Here is an example of how ranks differ among algorithms and field sizes using the word 'country':

To start showing the effects of the algorithms, the following example searches
the word 'country' using `query expansion`. This means that we are not looking
only the exact matches, but also the entries that appear the most when the
the exact match has been found.

In the `SELECT` clause you'll see both FTS expressions using `NATURAL LANGUAGE`
with query expansion and `BOOLEAN`modes respectively.

```
set global innodb_ft_aux_table = 'test/bookContentByLine';

SELECT  content, bookid, group_concat(it.POSITION) as pos,
        round(MATCH(content) AGAINST ("country" IN NATURAL LANGUAGE MODE WITH QUERY EXPANSION)) as QERank,
        round(MATCH(content) AGAINST ("country" IN BOOLEAN MODE)) as BoolRank
      FROM bookContentByLine bl join information_schema.INNODB_FT_INDEX_TABLE it
          ON (bl.FTS_DOC_ID = it.DOC_ID)  
      WHERE  MATCH(content) AGAINST ("country" IN NATURAL LANGUAGE MODE WITH QUERY EXPANSION)
          AND it.WORD = 'country'
     GROUP BY FTS_DOC_ID
     ORDER BY 4 DESC
     LIMIT 10 ;

+-------------------------------------------------------------------------------+--------+------+--------+----------+
| content                                                                       | bookid | pos  | QERank | BoolRank |
+-------------------------------------------------------------------------------+--------+------+--------+----------+
| "country in September, migrating into Holland, and leave their mates behind"  |  15707 | 1    |    105 |        7 |
| "unsatisfied desire to serve his country, the two prevalent enthusiasms at"   |  15707 | 33   |     98 |        7 |
| "Language, Vol. I. p. 212. In this country, where four or five horses travel" |  15707 | 35   |     93 |        7 |
| "inflicting immense damage upon the country. Whereupon the Florentines"       |   1232 | 36   |     89 |        7 |
| "made for a country of twenty or thirty millions’ population, can be laid"  |  39064 | 12   |     89 |        7 |
| "The spiders of this country manufacture nets of various forms, adapted to"   |  15707 | 21   |     87 |        7 |
| "a velvet-covered arm-chair at my head! This country is too decadent"         |  33415 | 45   |     86 |        7 |
| "country may be ennobled, and under its auspices may be verified that"        |   1232 | 1    |     84 |        7 |
| "name. The writer of this unpublished pamphlet sees his country in a"         |  39064 | 56   |     84 |        7 |
| "In our country, Mr. Pennant informs us, that some quails migrate, and"       |  15707 | 8    |     83 |        7 |
| "all the morning in passing over the adjacent country." (Voyage to Senegal,"  |  15707 | 46   |     82 |        7 |
| "the electoral system of the country. Immediately an outcry burst out"        |  39064 | 29   |     82 |        7 |
| "country, under a most excellent president, wherein all cities had their"     |   1232 | 1    |     81 |        7 |
| "Though in this country horses shew little vestiges of policy, yet in the"    |  15707 | 16   |     81 |        7 |
| "country districts. As Lucca had five gates, he divided his own country"      |   1232 | 1,63 |     80 |       14 |
+-------------------------------------------------------------------------------+--------+------+--------+----------+
15 rows in set (1,16 sec)

```

The noise generated by the query expansion is expected and described in the official documentation [here.](https://dev.mysql.com/doc/refman/5.7/en/fulltext-query-expansion.html)

The interesting case is the following row, which has 2 exact occurrences and it is not the highest rank using query
extension. Remember, this is expected.

```
| "country districts. As Lucca had five gates, he divided his own country" |   1232 | 1,63 |     80 |       14 |
```

This is even worser when using large sentences. In the example bellow you will see the same query, against the table
storing by paragraph. The boolean rank shows some of the entries way above others, however the query extension
locates at the top records that not necessarily has a lot of exact matches.

```

(test) > set global innodb_ft_aux_table = 'test/bookContent';
Query OK, 0 rows affected (0,00 sec)

(test) > SELECT bookid, FTS_DOC_ID,
                group_concat(it.POSITION) as positions,
                round(MATCH(content) AGAINST ("country" IN NATURAL LANGUAGE MODE WITH QUERY EXPANSION)) as QERank,
                round(MATCH(content) AGAINST ("country" IN BOOLEAN MODE)) as BooleanRank,
                length(content) as len
          FROM bookContent bl join information_schema.INNODB_FT_INDEX_TABLE it ON (bl.FTS_DOC_ID = it.DOC_ID)  
          WHERE  MATCH(content) AGAINST ("country" IN NATURAL LANGUAGE MODE WITH QUERY EXPANSION)
              AND it.WORD = 'country'
          GROUP BY FTS_DOC_ID
          ORDER BY QERank DESC
          LIMIT 10 ;

+--------+------------+-----------------+--------+-------------+-------+
| bookid | FTS_DOC_ID | positions       | QERank | BooleanRank | len   |
+--------+------------+-----------------+--------+-------------+-------+
|  16452 |      17637 | 942,2552,9084   |  32494 |          10 | 51790 |
|  16452 |      17827 | 31699           |  30232 |           3 | 51701 |
|  16452 |      17761 | 667,47646       |  29517 |           7 | 50264 |
|  16452 |      17791 | 13566           |  28888 |           3 | 49129 |
|  16452 |      17927 | 23259,7044      |  26731 |           7 | 48983 |
|  16452 |      17839 | 9012,199        |  24933 |           7 | 44451 |
|  16452 |      17815 | 29318           |  24745 |           3 | 44011 |
|  16452 |      17729 | 895,16485,24034 |  23305 |          10 | 42612 |
|  16452 |      17621 | 1765            |  19935 |           3 | 36698 |
|  16452 |      17803 | 3942            |  17552 |           3 | 30586 |
+--------+------------+-----------------+--------+-------------+-------+
10 rows in set (1,88 sec)
```

```{r, message=FALSE, echo=FALSE}
sql <- "SELECT avg(length(content)) as len from bookContent"
rs <- dbGetQuery(con,sql)
```

The query expansion is useful when you intend to search which entries contain
more words that appear frequently within the search term. Having large text fields
increase the probability to have more words that appear among the search term.
In the case of `bookContent` table (by paragraph table), the average field size
is `r rs$len` characters.

## The INNODB_FT_INDEX_TABLE

There is a way to play with the contents of the FTS indexes. As you ma noticed
in the previos examples, I used the `set global innodb_ft_aux_table = 'test/bookContent';`
statement, which loads the index content to memory for an easy querying.

i.e. You can easily get the most frequent tokens:

```
(test) > select WORD,count(*)
           FROM information_schema.INNODB_FT_INDEX_TABLE   
           group by WORD having count(*) > 1000
           order by 2
           limit 10;
+--------+----------+
| WORD   | count(*) |
+--------+----------+
| should |     1023 |
| yet    |     1027 |
| any    |     1070 |
| like   |     1071 |
| been   |     1073 |
| first  |     1080 |
| nor    |     1087 |
| your   |     1106 |
| thou   |     1130 |
| shall  |     1164 |
+--------+----------+
10 rows in set (5,40 sec)
```

Probably it isn't a very useful information as most of this words appears too frequently
and are modal verbs,  adverbs,  pronouns, determiners, etc.

It could be the case that you are not interested on indexing those words. If that's
the case you can add them as `stopwords` in your own stopwords table. I built I
query for this situation to allow us to build the stopwords table using the current
words that we want to add to the filtering:

```
(test) > select group_concat(WORD) FROM (select distinct WORD  
  FROM information_schema.INNODB_FT_INDEX_TABLE               
  group by WORD having count(*) > 1000) d\G
*************************** 1. row ***************************
group_concat(WORD): all,and,any,been,but,can,first,had,has,have,her,him,his,into,
its,like,may,more,nor,not,now,one,only,other,our,said,shall,she,should,some,such,
than,their,them,then,there,these,they,those,thou,thus,thy,time,were,which,would,
yet,you,your
1 row in set (5,28 sec)

```

Let's build our filter table using both default and new entries and keeping the
alphabetical order:

```
CREATE TABLE bookContentByLine_stopwords(value VARCHAR(30)) ENGINE = INNODB;

INSERT INTO bookContentByLine_stopwords
SELECT value FROM (
    SELECT value FROM
    INFORMATION_SCHEMA.INNODB_FT_DEFAULT_STOPWORD
    UNION
    SELECT DISTINCT WORD as value  
      FROM information_schema.INNODB_FT_INDEX_TABLE               
      GROUP BY WORD having count(*) > 1000
) allEntries
ORDER BY value ASC;

DROP INDEX ftscontent ON bookContentByLine;
SET GLOBAL innodb_ft_server_stopword_table = 'test/bookContentByLine_stopwords';
CREATE FULLTEXT INDEX ftscontent ON bookContentByLine(content);

```

TODO: Check index sizes

```
(test) > select *
          from information_schema.INNODB_FT_INDEX_TABLE
          WHERE WORD = 'country';
+------------------+--------------+-------------+-----------+--------+----------+
| WORD             | FIRST_DOC_ID | LAST_DOC_ID | DOC_COUNT | DOC_ID | POSITION |
+------------------+--------------+-------------+-----------+--------+----------+
| country          |          149 |         787 |        28 |    733 |      265 |
| country          |          149 |         787 |        28 |    733 |     1342 |
| countrydistricts |          733 |         733 |         1 |    733 |      816 |
| thecountry       |          249 |         733 |         2 |    733 |      750 |
+------------------+--------------+-------------+-----------+--------+----------+
4 rows in set (0,08 sec)

(test) > select *
          from information_schema.INNODB_FT_INDEX_TABLE
          WHERE DOC_ID = 155 AND lower(WORD) like '%country%';
+---------+--------------+-------------+-----------+--------+----------+
| WORD    | FIRST_DOC_ID | LAST_DOC_ID | DOC_COUNT | DOC_ID | POSITION |
+---------+--------------+-------------+-----------+--------+----------+
| country |          149 |         787 |        28 |    155 |       31 |
| country |          149 |         787 |        28 |    155 |      495 |
| country |          149 |         787 |        28 |    155 |      158 |
| country |          149 |         787 |        28 |    155 |      525 |
+---------+--------------+-------------+-----------+--------+----------+
4 rows in set (0,09 sec)

```

> In the example shown before the is no intention to compare ranks score as they are based in different algorithms.
> The idea there is to show that QUERY EXPANSION can have non desire results in some cases due to its mechanism.

## Considerations if you are planning to use Innodb's Full text search

- Use QUERY EXPANSION only if you are interested to search relations over exact matches. Remember that the field
  size is crucial when using this.
- FTS is not the best fit for exact string matches. You won't use FTS for searching emails or strings as words using
  single fields.
- If you are using BOOLEAN MODE, you can use the rank score to filter rows. MySQL is clever enough to optimize the
  FTS functions to avoid double executions. You can do this using somehting like:
  `match(content,title) against ("first (<second >third)") > 1 `
  Generally, scores lower than 1 can be ignored when using boolean or natural mode searches.
- `OPTIMIZE TABLE` does a rebuild of the table. To avoid this, set `innodb_optimize_fulltext_only=1` in order to do an incremental
  maintance on the table.
- Recall that NATURAL LANGUAGE MODE does not take the operands as the BOOLEAN MODE.
- If you plan to order by rank, it is not necessary to specify the clause `ORDER BY` as MySQL does the order forcelly internally. Also,
  the behavior is different from the default as it returns the heaviest at the top (like ORDER BY rank DESC).
- If you come from MyISAM's FTS implementation, recall that the ranking scoring is different.
- Create the FULLTEXT index after the data is loaded [InnoDB bulk load](http://dev.mysql.com/doc/refman/5.7/en/optimizing-innodb-bulk-data-loading.html). When restoring FTS backups, you will probably hit the "ERROR 182 (HY000) at line nn: Invalid InnoDB FTS Doc ID".
- Try to don't use more than one FTS expression in the where clause. Keep in mind that this affects the order in the results and
  it consumes a considerably amount of CPU. MySQL orders by the latest expression in the WHERE clause. [WL#7123](https://dev.mysql.com/worklog/task/?id=7123)
- Also, if avoiding the rank information in the projection (SELECT clause) and using other aggregations like `count(*)`, will use
  the "no ranking" FT_hints. The `limit` hint won't be used if invoked explicitely an `ORDER BY` and th `MATCH` clause in the projection.

```
(test) >  explain  select  *
                      from bookContentByLine
                      where match(content) against ("+home" IN BOOLEAN MODE)
                      ORDER BY FTS_DOC_ID  
                      LIMIT 10\G
  select_type: SIMPLE
        table: bookContentByLine
         type: fulltext
        Extra: Using where; Ft_hints: no_ranking; Using filesort

(test) >  explain  select  *
                      from bookContentByLine
                      where match(content) against ("+home" IN BOOLEAN MODE)
                      LIMIT 10\G
        table: bookContentByLine
         type: fulltext
        Extra: Using where; Ft_hints: no_ranking, limit = 10

(test) >  explain  select  count(content)
                    from bookContentByLine
                    where match(content) against ("+home" IN BOOLEAN MODE) \G
        table: bookContentByLine
         type: fulltext
        Extra: Using where; Ft_hints: no_ranking

```
- If you plan to use FTS_DOC_ID column with AUTO_INCREMENT option, have in mind that there is a limitation regarding this.You must
  declare a single column PRIMARY KEY constraint or as an UNIQUE index. Also, the data type is stricted as `bigint unsigned`. I.e:

```
CREATE TABLE test ( FTS_DOC_ID bigint unsigned  AUTO_INCREMENT,
                    mainPk bigint, other text,
                    PRIMARY KEY(mainPk),
                    UNIQUE(FTS_DOC_ID)
                    );
```

## Stats of the datasets

Here we prompt some stats regarding the used datasets for the tests and data written here:

```{r, echo=FALSE}

sql <- "SELECT bp.bookid, round(avg(length(bp.content))) as avgLength, round(STD(length(bp.content))) as stdLength, count(*) as numEntries
            FROM bookContent bp  
            GROUP BY bookid;"
rs <- dbGetQuery(con, sql)
return(rs)

bookContentStats <- rs

sql <- "SELECT bl.bookid, round(avg(length(bl.content))) as avgLength, round(STD(length(bl.content))) as stdLength, count(*) as numEntries
            FROM  bookContentByLine bl
            GROUP BY bookid;"
rs <- dbGetQuery(con, sql)
return(rs)

bookContentByLineStats <- rs

```

## Playing with searches

The following form will allow you to search across the stats over the tables of the filtering provided dynamically.

The aggregations shown on the results are over the results using the selected algorithm.

You can start trying `+country +(>house <president) -home` in boolean mode. For more information about operator in Innodb's FTS,
you can read [this.](https://dev.mysql.com/doc/refman/5.7/en/fulltext-boolean.html) Using boolean formulas in NATURAL LANGUAGE MODE,
does not have the desired effect, however you'll be able to see how it affects the ranking compared with the boolean search.

NOTE: distance operator not yet supported.

```{r, echo=FALSE}
        ui <- shinyUI(fluidPage(
                mainPanel(  
                          selectInput("ftsMode", label = "Algorithm:",
                                      choices = c("WITH QUERY EXPANSION", "IN BOOLEAN MODE", "IN NATURAL LANGUAGE MODE"), selected = "WITH QUERY EXPANSION"),
                          textInput("filtering", "FTS Filter:", "default"),
                          submitButton("Stats!"),
                          # uiOutput('tableResults'),
                          uiOutput('table'),
                          uiOutput('tableResults')
                          )
                       ))

        server <- function(input, output, session) {

                DataResults = reactive({
                   f<-function(x) {as.character(input$filtering)}
                  sql <- paste("SELECT FTS_DOC_ID, bookid,
                              MATCH(content) AGAINST(\"" ,f(x)," \" ",input$ftsMode,"), content
                              FROM bookContentByLine WHERE MATCH(content) AGAINST(\"" ,f(x)," \" ",input$ftsMode,") LIMIT 50", sep = "")

                  rs <- dbGetQuery(con, sql)
                  return(rs)

                })

                output$tableResults <- renderTable({DataResults()})

                Data = reactive({
                  f<-function(x) {as.character(input$filtering)}

                      sql <- paste("SELECT sourceTable,
                                      max(scoreBoolean) as maxBoolean,
                                      max(scoreQE) as maxQE,
                                      -- max(scoreNL) as maxNL,
                                      avg(scoreBoolean) as avgBoolean,
                                      avg(scoreQE) as avgQE,
                                      -- avg(scoreNL) as avgNL,
                                      avg(lengthContent) avgFieldLength,
                                      -- avg(scoreQE) / count(sourceTable) as weight,
                                      count(sourceTable) as countOccurrences
                              FROM
                              (
                              SELECT \"byParagraph\" as sourceTable,
                                    FTS_DOC_ID,
                                    match(BC.content) against (\"" ,f(x) ," \" IN BOOLEAN MODE) as scoreBoolean,
                                    match(BC.content)
                                        against (\""  ,f(x), "\" IN NATURAL LANGUAGE MODE WITH QUERY EXPANSION) scoreQE,
                                    match(BC.content)
                                        against (\""  ,f(x), "\" IN NATURAL LANGUAGE MODE) scoreNL,
                                     length(content) as lengthContent
                              FROM    bookContent BC
                              WHERE
                                match(BC.content) against (\"" ,f(x)," \" ",input$ftsMode,")
                              UNION
                              SELECT \"byLine\" as sourceTable,
                                      FTS_DOC_ID, match(BL.content) against (\"",f(x)," \" IN BOOLEAN MODE) as scoreBoolean,
                                     match(BL.content)
                                      against (\"" ,f(x)," \" IN NATURAL LANGUAGE MODE WITH QUERY EXPANSION) scoreQE,
                                    match(BL.content)
                                        against (\""  ,f(x), "\" IN NATURAL LANGUAGE MODE) scoreNL,
                                     length(content) as lengthContent
                              FROM
                                      bookContentByLine BL
                              WHERE
                                match(BL.content) against (\"", f(x) ,"\" ",input$ftsMode,")
                              ) unionTable
                              GROUP BY sourceTable
                              ;" , sep = "")

                      rs <- dbGetQuery(con, sql)
                      return(rs)

                })

                output$table <- renderTable({Data()})

        }

shinyApp(ui = ui, server = server)

```

Now that you can see the general stats, we'll see how much noise has been added with the query extension algorithm.
To do that,  we'll gather all the occurrences using the query extension option and extract only the ones that match
with the occurrence.

### Fine tuning

[Fine tuning](https://dev.mysql.com/doc/refman/5.7/en/fulltext-fine-tuning.html)

### Features introduces

### Maintenance

[innodb_optimize_fulltext_only](https://dev.mysql.com/doc/refman/5.7/en/innodb-parameters.html#sysvar_innodb_optimize_fulltext_only)

### Parsers internals

[Writting FTS parser plugins](https://dev.mysql.com/doc/refman/5.7/en/writing-full-text-plugins.html)