### 2016-04-20 article.rmd
Fixes

---
title: "InnoDB's Full Text Search overview (with Shiny/R)"
author: "Emanuel Calvo - Pythian"
date: "January 3rd, 2016"
output: html_document
runtime: shiny
---

```{r init, message = FALSE, echo=FALSE,include=FALSE}
library(png)
#options(width = 9999)
#knitr::opts_chunk$set(fig.width=9999)
# opts_chunk$set(fig.width=9999) #, fig.height=6))

library(knitr)
library(shiny)
suppressWarnings(library(RMySQL))
library(ggplot2)

# Comment if you want to avoid reload all the data
# If you want to start from scratch, uncomment the following line.
# Otherwise, you can restore the dumps located in the dumps folder

#source('~/fts_article/load.R')

mypass <- scan(file = '.key', what = (pass=""))
RDShost <- scan(file = '.RDShost', what = (RDShost=""))

# Yeah, only for development environements
all_cons <- dbListConnections(MySQL())
for(con in all_cons) dbDisconnect(con)

# You do not want clear passwords, but is a sandbox.
con <- dbConnect(MySQL(),
                 user="mysql57", password=mypass, port=3306
                 ,dbname="ftslab", 
                 host=RDShost)

```

## Credits

> Author: Emanuel Calvo
> Company: Pythian
>
> Thanks to Valerie Parham-Thompson @ Pythian and Daniel Prince @ Oracle.
>
> Repository available at [Github](https://github.com/3manuek/fts_article).

## Requirements

If you want to run this locally, please install a MySQL 5.7 instance and:

- Uncomment the line __source('~/fts_article/load.R')__ and run the report, or
- Restore the dumps located in the `dumps` folder. Uncomment the table creation.

## Some initial thoughts

A couple of days ago one of our customers came up with a question regarding FTS
over InnoDB engine. Although the question is not answered in the current article,
I came up with the conclusion that FTS is sometimes misunderstood.

The point of this _article_ is to show dynamically how the search algorithms work,
using non-fictional data (data sources were downloaded from www.gutenberg.org
project) within an easy interface.

In order to show the effects off the field sizes over the query expansion algorithm,
you will see two main tables (bookContent and bookContentByLine) both containing
the same books in different approaches: by line and by paragraph. You'll see the
noise generated by the `QUERY EXPANSION` algorithm when phrases are too large.

The current article has been developed using Shiny/R in order to allow you to see
the effects of the algorithms.

For the sake of simplicity, in this article we won't go through the FTS parsers.
Probably that would be material for a future post.

## Why I consider FTS sometimes misunderstood?

FTS is a technology that can be use for any purpose, not only simple searches.
There is a myth that the FTS only should be placed on clusters for that purpose,
which I agree. However, certain bussiness rules require comple searches, and 
having such feature can be a win.

RDBMS aren't a good place for massive amount of FTS queries, without using any
of the join capabilities that they offer, or the ACID complaints.

As I said above, FTS is totally acceptable in RDBMS, if you are using at least
one RDBMS critical feature, required by your bussines model.

## Action!

Here is an example of how ranks differ among algorithms and field sizes using the word 'country':

To start showing the effects of the algorithms, the following example searches
the word 'country' using `query expansion`. This means that we are not looking
only the exact matches, but also the entries that appear the most when the
the exact match has been found.

In the `SELECT` clause you'll see both FTS expressions using `NATURAL LANGUAGE`
with query expansion and `BOOLEAN`modes respectively.

```
set global innodb_ft_aux_table = 'ftslab/bookContentByLine';

SELECT  content, bookid, group_concat(it.POSITION) as pos,
        round(MATCH(content) AGAINST ("country" IN NATURAL LANGUAGE MODE WITH QUERY EXPANSION)) as QERank,
        round(MATCH(content) AGAINST ("country" IN BOOLEAN MODE)) as BoolRank
      FROM bookContentByLine bl join information_schema.INNODB_FT_INDEX_TABLE it
          ON (bl.FTS_DOC_ID = it.DOC_ID)  
      WHERE  MATCH(content) AGAINST ("country" IN NATURAL LANGUAGE MODE WITH QUERY EXPANSION)
          AND it.WORD = 'country'
     GROUP BY FTS_DOC_ID
     ORDER BY 4 DESC
     LIMIT 10 ;

+-----------------------------------------------------------------------------+--------+-----+--------+--------+
|content                                                                      | bookid | pos | QERank |BoolRank|
+-----------------------------------------------------------------------------+--------+-----+--------+--------+
|"country in September, migrating into Holland, and leave their mates behind" |  15707 | 1   |    105 |      7 |
|"unsatisfied desire to serve his country, the two prevalent enthusiasms at"  |  15707 | 33  |     98 |      7 |
|"Language, Vol. I. p. 212. In this country, where four or five horses travel"|  15707 | 35  |     93 |      7 |
|"inflicting immense damage upon the country. Whereupon the Florentines"      |   1232 | 36  |     89 |      7 |
|"made for a country of twenty or thirty millions’ population, can be laid" |  39064 | 12  |     89 |      7 |
|"The spiders of this country manufacture nets of various forms, adapted to"  |  15707 | 21  |     87 |      7 |
|"a velvet-covered arm-chair at my head! This country is too decadent"        |  33415 | 45  |     86 |      7 |
|"country may be ennobled, and under its auspices may be verified that"       |   1232 | 1   |     84 |      7 |
|"name. The writer of this unpublished pamphlet sees his country in a"        |  39064 | 56  |     84 |      7 |
|"In our country, Mr. Pennant informs us, that some quails migrate, and"      |  15707 | 8   |     83 |      7 |
|"all the morning in passing over the adjacent country." (Voyage to Senegal," |  15707 | 46  |     82 |      7 |
|"the electoral system of the country. Immediately an outcry burst out"       |  39064 | 29  |     82 |      7 |
|"country, under a most excellent president, wherein all cities had their"    |   1232 | 1   |     81 |      7 |
|"Though in this country horses shew little vestiges of policy, yet in the"   |  15707 | 16  |     81 |      7 |
|"country districts. As Lucca had five gates, he divided his own country"     |   1232 | 1,63|     80 |     14 |
+-----------------------------------------------------------------------------+--------+-----+--------+--------+
15 rows in set (1,16 sec)

```

The noise generated by the query expansion is expected and described in the official documentation [here](https://dev.mysql.com/doc/refman/5.7/en/fulltext-query-expansion.html).

The interesting case is the following row, which has 2 exact occurrences and it is not the highest rank using query
extension. Remember, this is expected.

```
| "country districts. As Lucca had five gates, he divided his own country" |   1232 | 1,63 |     80 |       14 |
```

This is even worser when using large sentences. In the example bellow you will see the same query, against the table
storing by paragraph. The boolean rank shows some of the entries way above others, however the query extension
locates at the top records that not necessarily has a lot of exact matches.

```

(ftslab) > set global innodb_ft_aux_table = 'ftslab/bookContent';
Query OK, 0 rows affected (0,00 sec)

(ftslab) > SELECT bookid, FTS_DOC_ID,
                group_concat(it.POSITION) as positions,
                round(MATCH(content) AGAINST ("country" IN NATURAL LANGUAGE MODE WITH QUERY EXPANSION)) as QERank,
                round(MATCH(content) AGAINST ("country" IN BOOLEAN MODE)) as BooleanRank,
                length(content) as len
          FROM bookContent bl join information_schema.INNODB_FT_INDEX_TABLE it ON (bl.FTS_DOC_ID = it.DOC_ID)  
          WHERE  MATCH(content) AGAINST ("country" IN NATURAL LANGUAGE MODE WITH QUERY EXPANSION)
              AND it.WORD = 'country'
          GROUP BY FTS_DOC_ID
          ORDER BY QERank DESC
          LIMIT 10 ;

+--------+------------+-----------------+--------+-------------+-------+
| bookid | FTS_DOC_ID | positions       | QERank | BooleanRank | len   |
+--------+------------+-----------------+--------+-------------+-------+
|  16452 |      17637 | 942,2552,9084   |  32494 |          10 | 51790 |
|  16452 |      17827 | 31699           |  30232 |           3 | 51701 |
|  16452 |      17761 | 667,47646       |  29517 |           7 | 50264 |
|  16452 |      17791 | 13566           |  28888 |           3 | 49129 |
|  16452 |      17927 | 23259,7044      |  26731 |           7 | 48983 |
|  16452 |      17839 | 9012,199        |  24933 |           7 | 44451 |
|  16452 |      17815 | 29318           |  24745 |           3 | 44011 |
|  16452 |      17729 | 895,16485,24034 |  23305 |          10 | 42612 |
|  16452 |      17621 | 1765            |  19935 |           3 | 36698 |
|  16452 |      17803 | 3942            |  17552 |           3 | 30586 |
+--------+------------+-----------------+--------+-------------+-------+
10 rows in set (1,88 sec)
```

```{r, message=FALSE, echo=FALSE}
sql <- "SELECT avg(length(content)) as len from bookContent"
rs <- dbGetQuery(con,sql)
```

The query expansion is useful when you intend to search which entries contain
more words that appear frequently within the search term. Having large text fields
increase the probability to have more words that appear among the search term.
In the case of `bookContent` table (by paragraph table), the average field size
is `r rs$len` characters.

## The `INNODB_FT_INDEX_TABLE`

There is a way to play with the contents of the FTS indexes. As you may noticed
in the previous examples, I used the `set global innodb_ft_aux_table = 'ftslab/bookContent';`
statement, which loads the index content to memory for an easy querying.

If you use RDS, the option `innodb_ft_aux_table` is not available as it is GLOBAL
and require SUPER privileges.

i.e. You can easily get the most frequent tokens:

```
(ftslab) > SELECT WORD,count(*)
           FROM information_schema.INNODB_FT_INDEX_TABLE   
           group by WORD having count(*) > 1000
           order by 2
           limit 10;
+--------+----------+
| WORD   | count(*) |
+--------+----------+
| should |     1023 |
| yet    |     1027 |
| any    |     1070 |
| like   |     1071 |
| been   |     1073 |
| first  |     1080 |
| nor    |     1087 |
| your   |     1106 |
| thou   |     1130 |
| shall  |     1164 |
+--------+----------+
10 rows in set (5,40 sec)
```

Probably it isn't a very useful information as most of this words appears too frequently
and are modal verbs,  adverbs,  pronouns, determiners, etc. It could be the case that you
are not interested on indexing those words. If that's the case you can add them as `stopwords` 
in your own stopwords table. Specially if you are more interested in boolean searches, loosing
some part of the language expressions. I built I query for this situation to allow us to build 
the stopwords table using the current words that we want to add to the filtering:

```
(ftslab) > select group_concat(WORD) FROM (select distinct WORD  
  FROM information_schema.INNODB_FT_INDEX_TABLE               
  group by WORD having count(*) > 1000) d\G
*************************** 1. row ***************************
group_concat(WORD): all,and,any,been,but,can,first,had,has,have,her,him,his,into,
its,like,may,more,nor,not,now,one,only,other,our,said,shall,she,should,some,such,
than,their,them,then,there,these,they,those,thou,thus,thy,time,were,which,would,
yet,you,your
1 row in set (5,28 sec)

```

Let's build our filter table using both default and new entries and keeping the
alphabetical order:

```
CREATE TABLE bookContentByLine_stopwords(value VARCHAR(30)) ENGINE = INNODB;

INSERT INTO bookContentByLine_stopwords
SELECT value FROM (
    SELECT value FROM
    INFORMATION_SCHEMA.INNODB_FT_DEFAULT_STOPWORD
    UNION
    SELECT DISTINCT WORD as value  
      FROM information_schema.INNODB_FT_INDEX_TABLE               
      GROUP BY WORD having count(*) > 1000
) allEntries
ORDER BY value ASC;

DROP INDEX ftscontent ON bookContentByLine;
SET GLOBAL innodb_ft_server_stopword_table = 'ftslab/bookContentByLine_stopwords';
CREATE FULLTEXT INDEX ftscontent ON bookContentByLine(content);

```

Checking the contents of the index is easy as:

```
(ftslab) > select *
          from information_schema.INNODB_FT_INDEX_TABLE
          WHERE lower(WORD) like '%country%';
+------------------+--------------+-------------+-----------+--------+----------+
| WORD             | FIRST_DOC_ID | LAST_DOC_ID | DOC_COUNT | DOC_ID | POSITION |
+------------------+--------------+-------------+-----------+--------+----------+
| country          |          149 |         787 |        28 |    733 |      265 |
| country          |          149 |         787 |        28 |    733 |     1342 |
| countrydistricts |          733 |         733 |         1 |    733 |      816 |
| thecountry       |          249 |         733 |         2 |    733 |      750 |
+------------------+--------------+-------------+-----------+--------+----------+
4 rows in set (0,08 sec)

(ftslab) > select *
          from information_schema.INNODB_FT_INDEX_TABLE
          WHERE DOC_ID = 155 AND lower(WORD) like '%country%';
+---------+--------------+-------------+-----------+--------+----------+
| WORD    | FIRST_DOC_ID | LAST_DOC_ID | DOC_COUNT | DOC_ID | POSITION |
+---------+--------------+-------------+-----------+--------+----------+
| country |          149 |         787 |        28 |    155 |       31 |
| country |          149 |         787 |        28 |    155 |      495 |
| country |          149 |         787 |        28 |    155 |      158 |
| country |          149 |         787 |        28 |    155 |      525 |
+---------+--------------+-------------+-----------+--------+----------+
4 rows in set (0,09 sec)

```

> In the example shown before the is no intention to compare ranks score as they are based in different algorithms.
> The idea there is to show that QUERY EXPANSION can have non desire results in some cases due to its mechanism.

### Going ahead on choosing stop words

The full [article](https://en.wikipedia.org/wiki/Zipf's_law) is amazingly interesting. In a brief, it says that
the most frequent word will occur approximately twice as often as the second most frequent word, three times as 
often as the third most frequent word, and so on (rank-frequency distribution is an inverse relation).

The idea here is to measure how much index do we safe cutting those words that are extremely frequent and don't add
a necessary meaning to the search.

## Considerations and recommendations

- Use QUERY EXPANSION only if you are interested to search relations over exact matches. Remember that the field
  size is crucial when using this.
- FTS is not the best fit for exact string matches in single columns. You don't want to use FTS for searching emails in
  a single column, name and lastname fields , i.e. For those, you'll probably use other techniques as reverse searches or
  exact match operator (=).
- Keep your FTS indexes short. Do not add ALL the text columns. Parse first from your application the user search and 
  adapt the query. 
- If you are using BOOLEAN MODE, you can use the rank score to filter rows. MySQL is clever enough to optimize the
  FTS functions to avoid double executions. You can do this using something like:
  `match(content,title) against ("first (<second >third)") > 1 `
  Generally, scores lower than 1 can be ignored when using boolean or natural mode searches.
- `OPTIMIZE TABLE` does a rebuild of the table. To avoid this, set `innodb_optimize_fulltext_only=1` in order to do an incremental
  maintance on the table.
- Recall that NATURAL LANGUAGE MODE does not take the operands as the BOOLEAN MODE. This affects the ranking score (try __+bad (<feeling >thing)__ i.e.)
- If you plan to order by rank, it is not necessary to specify the clause `ORDER BY` as InnoDB does the order after retrieve the doc ids . Also,the behavior is different from the default as it returns the heaviest at the top (like an __ORDER BY rank DESC__).
- If you come from MyISAM's FTS implementation, recall that the ranking scoring is different.
- Create the FULLTEXT index after the data is loaded [InnoDB bulk load](http://dev.mysql.com/doc/refman/5.7/en/optimizing-innodb-bulk-data-loading.html). When restoring FTS backups, you will probably hit the "ERROR 182 (HY000) at line nn: Invalid InnoDB FTS Doc ID".
- Try to avoid using use more than one FTS expression in the where clause. Keep in mind that this affects the order in the results and
  it consumes a considerably amount of CPU. InnoDB orders by the latest expression in the WHERE clause. [WL#7123](https://dev.mysql.com/worklog/task/?id=7123)
- Also, if avoiding the rank information in the projection (SELECT clause) and using other aggregations like `count(*)`, will use
  the "no ranking" FT_hints. The `limit` hint won't be used if invoked explicitely an `ORDER BY` and th `MATCH` clause in the projection.

```
(ftslab) >  explain  select  *
                      from bookContentByLine
                      where match(content) against ("+home" IN BOOLEAN MODE)
                      ORDER BY FTS_DOC_ID  
                      LIMIT 10\G
  select_type: SIMPLE
        table: bookContentByLine
         type: fulltext
        Extra: Using where; Ft_hints: no_ranking; Using filesort

(ftslab) >  explain  select  *
                      from bookContentByLine
                      where match(content) against ("+home" IN BOOLEAN MODE)
                      LIMIT 10\G
        table: bookContentByLine
         type: fulltext
        Extra: Using where; Ft_hints: no_ranking, limit = 10

(ftslab) >  explain  select  count(content)
                    from bookContentByLine
                    where match(content) against ("+home" IN BOOLEAN MODE) \G
        table: bookContentByLine
         type: fulltext
        Extra: Using where; Ft_hints: no_ranking

```

- If you plan to use `FTS_DOC_ID` column with `AUTO_INCREMENT` option, 
  have in mind that there is a limitation regarding this. You must declare 
  a single column PRIMARY KEY constraint or as an UNIQUE index. Also, the data 
  type is stricted as `bigint unsigned`. i.e:

```
CREATE TABLE test ( FTS_DOC_ID bigint unsigned  AUTO_INCREMENT,
                    mainPk bigint, other text,
                    PRIMARY KEY(mainPk),
                    UNIQUE(FTS_DOC_ID)
                    );
```

## Stats of the datasets

Here we prompt some stats regarding the used datasets for the tests and data written here:

```{r, ignore=FALSE, echo=FALSE, message=FALSE}

sql <- "SELECT bp.bookid, round(avg(length(bp.content))) as avgLength, round(STD(length(bp.content))) as stdLength, count(*) as numEntries
            FROM bookContent bp  
            GROUP BY bookid;"
rs1 <- dbGetQuery(con, sql)

bookContentStats <- rs1

sql <- "SELECT bl.bookid, round(avg(length(bl.content))) as avgLength, round(STD(length(bl.content))) as stdLength, count(*) as numEntries
            FROM  bookContentByLine bl
            GROUP BY bookid;"
rs2 <- dbGetQuery(con, sql)

bookContentByLineStats <- rs2

```

```{r, echo=FALSE}
return(rs1)
return(rs2)

```

## Playing with searches

The following form will allow you to search across the stats over the tables of the filtering provided dynamically.

The aggregations shown on the results are over the results using the selected algorithm.

You can start trying `+country +(>house <president) -home` in boolean mode. For more information about operator in Innodb's FTS,
you can read [this.](https://dev.mysql.com/doc/refman/5.7/en/fulltext-boolean.html) Using boolean formulas in NATURAL LANGUAGE MODE,
does not have the desired effect, however you'll be able to see how it affects the ranking compared with the boolean search.

NOTE: distance operator not yet supported.

```{r, echo=FALSE}
        ui <- shinyUI(fluidPage(
                mainPanel(  
                          selectInput("ftsMode", label = "Algorithm:",
                                      choices = c("WITH QUERY EXPANSION", "IN BOOLEAN MODE", "IN NATURAL LANGUAGE MODE"), selected = "WITH QUERY EXPANSION"),
                          textInput("filtering", "FTS Filter:", "default"),
                          submitButton("Stats!"),
                          # uiOutput('tableResults'),
                          uiOutput('table'),
                          uiOutput('tableResults')
                          )
                       ))

        server <- function(input, output, session) {

                DataResults = reactive({
                   f<-function(x) {as.character(input$filtering)}
                  sql <- paste("SELECT FTS_DOC_ID, bookid,
                              MATCH(content) AGAINST(\"" ,f(x)," \" ",input$ftsMode,") as rankScore, content
                              FROM bookContentByLine WHERE MATCH(content) AGAINST(\"" ,f(x)," \" ",input$ftsMode,") LIMIT 50", sep = "")

                  rs <- dbGetQuery(con, sql)
                  return(rs)

                })

                output$tableResults <- renderTable({DataResults()})

                Data = reactive({
                  f<-function(x) {as.character(input$filtering)}

                      sql <- paste("SELECT sourceTable,
                                      max(scoreBoolean) as maxBoolean,
                                      max(scoreQE) as maxQE,
                                      max(scoreNL) as maxNL,
                                      avg(scoreBoolean) as avgBoolean,
                                      avg(scoreQE) as avgQE,
                                      avg(scoreNL) as avgNL,
                                      avg(lengthContent) avgFieldLength,
                                      -- avg(scoreQE) / count(sourceTable) as weight,
                                      count(sourceTable) as countOccurrences
                              FROM
                              (
                              SELECT \"byParagraph\" as sourceTable,
                                    FTS_DOC_ID,
                                    match(BC.content) against (\"" ,f(x) ," \" IN BOOLEAN MODE) as scoreBoolean,
                                    match(BC.content)
                                        against (\""  ,f(x), "\" IN NATURAL LANGUAGE MODE WITH QUERY EXPANSION) scoreQE,
                                    match(BC.content)
                                        against (\""  ,f(x), "\" IN NATURAL LANGUAGE MODE) scoreNL,
                                     length(content) as lengthContent
                              FROM    bookContent BC
                              WHERE
                                match(BC.content) against (\"" ,f(x)," \" ",input$ftsMode,")
                              UNION
                              SELECT \"byLine\" as sourceTable,
                                      FTS_DOC_ID, match(BL.content) against (\"",f(x)," \" IN BOOLEAN MODE) as scoreBoolean,
                                     match(BL.content)
                                      against (\"" ,f(x)," \" IN NATURAL LANGUAGE MODE WITH QUERY EXPANSION) scoreQE,
                                    match(BL.content) 
                                        against (\""  ,f(x), "\" IN NATURAL LANGUAGE MODE) scoreNL,
                                     length(content) as lengthContent
                              FROM
                                      bookContentByLine BL
                              WHERE
                                match(BL.content) against (\"", f(x) ,"\" ",input$ftsMode,")
                              ) unionTable
                              GROUP BY sourceTable
                              ;" , sep = "")

                      rs <- dbGetQuery(con, sql)
                      return(rs)

                })

                output$table <- renderTable({Data()})

        }

shinyApp(ui = ui, server = server)

```

Now that you can see the general stats, we'll see how much noise has been added with the query extension algorithm.
To do that,  we'll gather all the occurrences using the query extension option and extract only the ones that match
with the occurrence.

### FT_QUERY_EXPANSION_LIMIT

This variable controls the number of top matches when using `WITH QUERY EXPANSION` (affects only MyISAM).

[reference](http://dev.mysql.com/doc/refman/5.7/en/server-system-variables.html#sysvar_ft_query_expansion_limit)

### Bug 80347 - Invalid InnoDB FTS Doc ID

> Submitted https://bugs.mysql.com/bug.php?id=80347

```
emanuel@3laptop ~/sandboxes/rsandbox_5_7_9 $ ./m dumpTest < full.dump 
ERROR 182 (HY000) at line 73: Invalid InnoDB FTS Doc ID

emanuel@3laptop ~/sandboxes/rsandbox_5_7_9 $ ./m dumpTest < ddl.dump 
emanuel@3laptop ~/sandboxes/rsandbox_5_7_9 $ ./m dumpTest < onlyData.dump 
emanuel@3laptop ~/sandboxes/rsandbox_5_7_9 $ ./m dumpTest < full.dump 
ERROR 182 (HY000) at line 73: Invalid InnoDB FTS Doc ID
```

mysqldump is not very clever if you use `FTS_DOC_ID`:

```
2016-02-13T22:11:53.125300Z 19 [ERROR] InnoDB: Doc ID 10002 is too big. Its difference with largest used Doc ID 1 cannot exceed or equal to 10000
```

It takes dumps without considering the restriction coded in 
`innobase/row/row0mysql.cc`:

```
Difference between Doc IDs are restricted within
4 bytes integer. See fts_get_encoded_len() 
```

The fix to this is backuping the table by chunks of 10000 documents.

### Fine tuning

[Fine tuning](https://dev.mysql.com/doc/refman/5.7/en/fulltext-fine-tuning.html)
[Performance](https://blogs.oracle.com/mysqlinnodb/entry/innodb_full_text_search_performance)

### Features introduces

### Maintenance

[innodb_optimize_fulltext_only](https://dev.mysql.com/doc/refman/5.7/en/innodb-parameters.html#sysvar_innodb_optimize_fulltext_only)

### Parsers internals

[Writting FTS parser plugins](https://dev.mysql.com/doc/refman/5.7/en/writing-full-text-plugins.html)

## Version

This R Markdown document is made interactive using Shiny. Unlike the more traditional
workflow of creating static reports, you can now create documents that allow your
readers to change the assumptions underlying your analysis and see the results immediately.  
To learn more, see [Interactive Documents](http://rmarkdown.rstudio.com/authoring_shiny.html).